A Guide to Pre-Processing High-Throughput Animal Tracking Data

### 2021-02-03 docs/ms_atlas_preproc_main_text_2021-02-03.tex
add caption tex update

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/fig_02_filtering_data.png}
    \caption{\textbf{(a)} A velocity-autocorrelated movement track simulated for 5,000 positions (black line) using the \texttt{smoove} package (Gurarie et al. 2017). Three kinds of errors have been artificially added: (1) each position (grey points) is offset from the canonical track with the addition of normally distributed small-scale error, (2) large-scale error has been added to 0.5\% of positions, and (3) 300 positions (indices 500 -- 800) have been displaced to the top-right of the track to simulate a gross distortion that affects a continuous subset of the track. 
    The goal of pre-processing such datasets is to get the estimated positions (grey points) to match the canonical track (black line) as closely as possible.  
    \textbf{(b)} Tracks can be quickly filtered by spatial bounds ($0.5 \leq Y \leq 1.0$; dashed grey lines) using the \texttt{atlastools} function  \texttt{atl\_filter\_bounds}.
    Setting the function argument \texttt{remove\_inside = FALSE} retains positions within user supplied bounds (dashed grey lines; green points), and excludes those outside (grey points).}
    \label{fig:figure_simulated_data}
\end{figure}

>>>>

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/fig_02_filtering_data.png}
    \caption{\textbf{(a)} A velocity-autocorrelated movement track simulated for 5,000 positions (black line) using the \texttt{smoove} package (Gurarie et al. 2017). Three kinds of errors have been artificially added: (1) each position (grey points) is offset from the canonical track with the addition of normally distributed small-scale error, (2) large-scale error has been added to 0.5\% of positions, and (3) 300 positions (indices 500 -- 800) have been displaced to the top-right of the track to simulate a gross distortion that affects a continuous subset of the track. 
    The goal of pre-processing such datasets is to get the estimated positions (grey points) to match the canonical track (black line) as closely as possible.  
    Tracks can be quickly filtered by spatial bounds ($0.5 \leq Y \leq 1.0$; dashed grey lines) using the \texttt{atlastools} function  \texttt{atl\_filter\_bounds}.
    Setting the function argument \texttt{remove\_inside = FALSE} retains positions within user supplied bounds (dashed grey lines; green points), and excludes those outside (grey points).
    Reducing large-scale outliers in a movement track. The goal is to remove positions far away from the canonical track (black line), and retain those with only reasonable displacements from the canonical track (green points). 
    \textbf{(b)} Location error may affect single observations resulting in point outliers or `spikes' (grey circles), but it may also affect continuous subsets of a track, called a `prolonged spike' (purple triangles). 
    The former may be targeted by filtering on appropriate covariates such as speed and turning angle using the \texttt{atlastools} function \texttt{atl\_filter\_covariates}.
    Here, we filter out positions with incoming and outgoing speeds $\geq$ the 90\textsuperscript{th} percentile (0.025) and turning angle $\geq$ the 10\textsuperscript{th} percentile (35$^{\circ}$).
    Howevever, prolonged spikes cannot be effectively corrected by targeting each coordinate pair in isolation.
    The \texttt{atlastools} function \texttt{atl\_remove\_reflections} to remove prolonged spikes (grey circles) in tracking data is an example of targeting track subsets with a common distortion (here, a geometric translation).
    While this method filters out the prolonged spike to return only positions with reasonable displacement from the canonical track (green points) in this example, conceptualising and implementing such algorithms is difficult. 
    Users are cautioned to frequently check this and similar methods' results.
    }
    \label{fig:figure_simulated_data}
\end{figure}
